---
title: "Course Project 1"
author: "KK Setshedi"
date: "30/09/2020"
output: html_document
---

# Loading library and reading data
```{r}
library(tidyverse)
data <- read.csv("activity.csv")
```
 
# Reprocessing of data
Change date format into date
```{r}
data$date <-  as.Date(data$date)
```


# Histogram without NA values
Histogram of total number of steps taken on each day is shown below:

```{r}
sum_steps<-aggregate(data$steps,by=list(data$date),FUN=sum,na.rm=TRUE) 

hist(sum_steps$x, 
      breaks=seq(from=0, to=25000, by=2500),
      col="Blue", 
      xlab="Total number of steps", 
      ylim=c(0, 20), 
      main="Histogram of the total number of steps taken each day\n(NA removed)")
```

# Mean and Median of Steps
```{r}
mean_steps <- mean(sum_steps$x)
median_steps <-  median(sum_steps$x)
```
The mean steps are `r mean_steps` and the median steps are `r median_steps`

# Time Series Plot
Time series plot of the average number of steps taken
```{r}
avg_steps <- aggregate(data$steps, by=list(data$interval), FUN=mean, na.rm=T)

colnames(avg_steps) <- c("interval","steps")

ggplot(aes(x=interval, y=steps), data=avg_steps)+geom_line()
```

# Maximum Average 5 minute interal

```{r}
fiveminave <- avg_steps[avg_steps$steps==max(avg_steps$steps), 1]
```
The 5-minute interval that, on average, contains the maximum number of steps is `r fiveminave`

# Imputing NA
## Here is cosw to describe and show a strategy for imputing missing data
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA's)
```{r}
missing <- sum(is.na(data$steps))
```
The total number of missing values in the dataset is `r missing`

Replace NA values with the mean of the steps

```{r}
data$steps[is.na(data$steps)] <- mean(data$steps, na.rm = T)
```
Here are some roes of new data set
```{r}
head(data)
```

# Histogram with Replaced NA values
Histogram of total number of steps taken on each day is shown below:
```{r}
sum_steps <- aggregate(data$steps, by=list(data$date), FUN=sum, na.rm = T)

hist(sum_steps$x, 
      breaks=seq(from=0, to=25000, by=2500),
      col="blue", 
      xlab="Total number of steps", 
      ylim=c(0, 30), 
      main="Total number of steps taken each day\n(NA replaced by mean)"
)
```

Mean and median number of steps taken each day after replacing NA values with mean
```{r}
 mean(sum_steps$x)
 median(sum_steps$x)
```

# Difference in activity patterns between weekdays and weekdends
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends 

```{r}
# convert date into weekdays
data$days = tolower(weekdays(data$date))

# Now categorised days into weekend and weekdays

data$day_type <- ifelse(data$days == "saturday" | data$days== "sunday","weekend","weekday")

# take mean steps taken on weekend or weekday in the intervals

avg_steps <- aggregate(data$steps, by=list(data$interval, data$day_type), FUN=mean,na.rm=T)

colnames(avg_steps) <- c("interval", "day_type", "steps")

ggplot(aes(x=interval,y=steps), data=avg_steps) +geom_line()+facet_wrap(~avg_steps$day_type)
```




